---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# validation

<!-- badges: start -->
<!-- badges: end -->

The goal of validation is to provide an easy approach for the documentation and implementation of package and function tests for the SCTO R validation project.

## Installation

You can install the development version of validation from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("aghaynes/validation")
```

## Testing

The package contains a `test` function which is used to run all tests for a named package.

```{r example}
library(validation)
## basic example code

test("presize")
```

The output of this function is a string which can be copied and pasted into an appropriate issue on the GitHub `pkg_validation` repository.


### Implementing new tests

All tests are stored in `inst/tests` in package specific folders. E.g. the `inst/tests/presize` folder contains tests relevant to the `presize` package.

(At least) three files are required. 

1. setup-`package`.R
2. info.txt 
3. test-`somename`.R

The `setup-package.R` file installs, updates and/or loads the package being tested and any other relevant steps (e.g. loading a dataset).

`info.txt` contains a plain text description of the tests.

test-`somename`.R contains the tests themselves. Tests should be written using `testthat` syntax, e.g.

```
test_that("'1:3' creates a sequence of 1, 2, 3", 
          expect_equal(1:3, c(1,2,3)))
```


